---
title: "Assignment 2"
subtitle: "Nicole Blackburn and Jianing Zou"
format: pdf
editor: visual
---

You may work in pairs or individually for this assignment. Make sure you join a group in Canvas if you are working in pairs. Turn in this assignment as an HTML or PDF file to ELMS. Make sure to include the R Markdown or Quarto file that was used to generate it.

```{r}
#| message: FALSE
#| label: Libraries
library(tidyverse)
library(epidatr)
library(censusapi)
```

In this assignment, you will pull from APIs to get data from various data sources and use your data wrangling skills to use them all together. You should turn in a report in PDF or HTML format that addresses all of the questions in this assignment, and describes the data that you pulled and analyzed. You do not need to include full introduction and conclusion sections like a full report, but you should make sure to answer the questions in paragraph form, and include all relevant tables and graphics.

Whenever possible, use piping and `dplyr`. Avoid hard-coding any numbers within the report as much as possible.

## Pulling from APIs

Our first data source is the Delphi COVIDcast data. You can access this using the Epidata API built by Carnegie Mellon University's Delphi Research group. Documentation for this API can be found here: <https://cmu-delphi.github.io/delphi-epidata/>. Here, we find the smoothed estimate of the proportion of people experiencing Covid-like symptoms by county from April 6, 2020 to April 14, 2020.

```{r}
#| label: covid-april
#| warning: false
covid_april <- pub_covidcast('fb-survey', 
                       'smoothed_wcli', 
                       'county', 
                       'day', 
                       time_values = c(20200406:20200414))
head(covid_april)
```

For more information about the data, see: <https://cmu-delphi.github.io/delphi-epidata/api/covidcast_signals.html>

Answer the following questions:

-   Change the data from long to wide format by including the estimate of Covid-like symptoms for each day as a column. There should be a column for `geo_value` as well as a column for each of the days in the dataset.

```{r}
#| label: april-wider
covid_april |>
  pivot_wider(id_cols = geo_value,
              names_from = time_value,
              values_from = value)
```

-   Find the mean, median, and variance of the estimate on each of the days from April 6, 2020 to April 14, 2020. (Note that this is not the appropriate way of finding the overall measures in reality because we aren't using weights)

```{r}
#| label: april-mean
covid_april |>
  group_by(time_value) |>
  summarize(
    mean = mean(value),
    median = median(value),
    variance = var(value)
  )
```

-   Which counties had the highest report Covid-like symptoms on each of the days within this range?

**Answer**

On April 6, 2020, county 36005 (Bronx County, NY) had the highest report of Covid-like symptoms. From April 7, 2020 - April 13, 2020, county 36087 (Rockland County, NY) has the highest report of Covid-like symptoms. On April 14, 2020, county 36079 (Putnam County, NY) had the highest report of Covid-like symptoms.

```{r}
#| label: april-highest-counties
covid_april |>
  select(time_value, geo_value, value) |>
  group_by(time_value) |>
  arrange(desc(value)) |>
  slice_head(n = 1)
```

Using the API, get the actual COVID cases from the JHU Cases and Deaths (using the link above, `confirmed_7dav_incidence_prop`) from May 6, 2020 to May 14, 2020. This is the number of confirmed COVID cases per 100,000 people. Find the correlation between reported COVID-like symptoms and actual COVID cases per 100,000 people within each county a month later. Is there a relationship?

**Answer**

The correlation coefficient is $0.08998326$, which indicates that there is a weak positive correlation between the April illness symptoms and may COVID cases by county. . The p-value of \<2.2e-16 indicates the correlation is statistically significant. The 95% condifidence interval is $[0.07078256, 0.10911729]$, which does not include zero, further indicating we can reject the null hypothesis that there is no correlation.

```{r}
#| label: covid-may
covid_may <- pub_covidcast('jhu-csse', 
                       'confirmed_7dav_incidence_prop', 
                       'county', 
                       'day', 
                       time_values = c(20200506:20200514))
```

```{r}
#| label: covid-combined
covid_both_apr <-
  covid_april |>
  select(geo_value, time_value, value) |>
  mutate(same_date = time_value + 30)

covid_both_may <-
  covid_may |>
  select(geo_value, time_value, value)

covid_both <- inner_join(covid_both_apr,
                         covid_both_may,
                         join_by(geo_value, same_date == time_value))
```

```{r}
#| label: cor-test
cor.test(covid_both$value.x, covid_both$value.y)
```

## Covidcast API Data + ACS

Now lets add another data set. The `censusapi` package provides a nice R interface for communicating with this API. However, before running queries we need an access key. This (easy) process can be completed here:

<https://api.census.gov/data/key_signup.html>

Once you have an access key, save it as a text file, then read this key in the `cs_key` object. We will use this object in all following API queries. Note that I called my text file `census-key.txt` – yours might be different!

```{r}
#| label: cs-key
#| eval: true
cs_key <- read_file("C:/Users/npbvb/OneDrive - University of Maryland/R_Projects/727/APIs/census-key")
```

```{r}
#| label: cs-key2
#| eval: false
cs_key <- read_file("census-key")
```

You can navigate through the documentation for all Census Data APIs here: <https://www.census.gov/data/developers/data-sets.html> Documentation for the 5-year ACS API can be found here: <https://www.census.gov/data/developers/data-sets/acs-5year.html>.

In the following, we request basic socio-demographic information (population, median age, median household income, income per capita) for counties in the US. The information about the variables used here can be found here: <https://api.census.gov/data/2022/acs/acs5/variables.html>.

```{r}
#| label: ACS-5year
acs <- getCensus(name = "acs/acs5",
                    vintage = 2020, 
                    vars = c("NAME", 
                             "B01001_001E", 
                             "B06002_001E", 
                             "B19013_001E", 
                             "B19301_001E"), 
                    region = "county",
                    key = cs_key)
head(acs)
```

Now, it might be useful to rename the socio-demographic variables (`B01001_001E` etc.) in our data set and assign more meaningful names.

```{r}
#| label: acs5-rename
acs <-
  acs %>%
  rename(pop = B01001_001E, 
         age = B06002_001E, 
         hh_income = B19013_001E, 
         income = B19301_001E)
```

It seems like we could try to use this location information listed above to merge this data set with the COVID data. However, we first have to clean the geography data to match the two datasets. The COVID data has a five digit geography code, with the first two digits representing the state and the last three representing the county within that state. The ACS data has this separated out. Add a new variable `location` to the ACS data that has the geography value in the same format as the COVID data.

```{r}
#| label: acs-location

acs <- acs |> 
        mutate(
        location = paste0(state, county),
         .after = county)
```

Answer the following questions with the COVID data and ACS data.

-   First, check how many counties aren't matched. Then, create a new data set by joining the two datasets. Keep only counties that appear in both data sets.

**Answer**

In the covid dataset, there were 51 counties that weren't matched to the ACS dataset. These counties ended with 000. Per the Delphi Epidata API on Geographic coding, "Megacounty estimates are reported with a FIPS code ending with 000, which is never a FIPS code for a real county". In the ACS dataset, there were 1,810 counties that didn't match to the covid dataset. After combining the dataset, there are 1,391 counties that appear in both data sets.

```{r}
#| label: join-covid-acs
#| message: false
#| results: hide
covid_both |> 
  distinct(location = geo_value) |> 
  anti_join(acs) 

acs |> 
  distinct(geo_value = location) |> 
  anti_join(covid_both) 

combined_full <- inner_join(acs, covid_both, join_by(location == geo_value)) 

combined_full |> 
  count(location) |> 
  filter(n > 1)
```

-   Compute the mean of the proportion of people with covid-like illness symptoms on April 6, 2020 for counties that have an above average median household income and for those that have an below average median household income. When building your pipe, start with creating the grouping variable and then proceed with the remaining tasks. What conclusions might you draw from this?

**Answer**

On April 6th, 2020, counties with an above average median household income had a mean proportion of $1.0148242$ covid-like illness symptoms, while counties with a below average median household income had a mean proportion of $0.8810505$ covid-like illness symptoms. A possible explanation of this is the exposure to COVID through a job; households with two or more working adults and a higher household income have more exposure to COVID through multiple workplaces while households with only one working adult and a lower household income have exposure to COVID from only one workplace. Another conclusion could be that counties with above average household incomes reported covid-like symptoms more than counties with below average household incomes.

```{r}
#| label: symptoms-mean-hhinc

combined <- combined_full |> 
  select(location, hh_income, time_value, value.x)

combined |>
  group_by(above_avg_hh = hh_income > mean(combined$hh_income)) |>
  filter(time_value == as.Date("2020-04-06")) |>
  summarize(mean_prop = mean(value.x))
```

-   Is there a relationship between the median household income and the proportion of people reporting Covid-like illness symptoms? Describe the relationship and use a scatterplot.

**Answer**

The correlation coefficient of $-0.06415457$ indicates a weak, negative relationship between median household income and the proportion of people reporting covid-like symptoms. The p-value of 1.763e-10 is less than 0.05, showing the relationship to be statistically significant. The 95% confidence interval $(-0.08377330, -0.04448613)$ does not include zero, further indicating that we can reject the null hypothesis of having no relationship.

```{r}
#| label: correlation-income
cor.test(combined$hh_income, combined$value.x)

ggplot(combined, aes(x = hh_income, y = value.x)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  labs(
    x = "Median Household Income",
    y = "Proportion reporting COVID-like symptoms",
    title = "COVID-like Illness vs. Median Household Income"
  )
```

## Using Other Census Data

Suppose we wanted to use the 2020 1-year ACS instead of the 5-year ACS. Why would we be unable to do this?

*Hint: Read the documentation for the 1-year ACS*

**Answer**

We can't do this because The Census Bureau does not release its standard 2020 ACS 1-year estimates because of the impact of the COVID-19 pandemic on data collection. Experimental estimates, developed from 2020 ACS 1-year data are available on the ACS Experimental Data page. They will not be available on data.census.gov or the Application Programming Interface (API).

Instead, repeat the steps above to merge the Delphi COVIDcast data to the 1-year ACS from 2021 (rather than the 5-year ACS). Do the same analysis as above.

```{r}
#| label: acs-1year

acs_2021 <- getCensus(name = "acs/acs1",
                    vintage = 2021, 
                    vars = c("NAME", 
                             "B01001_001E", 
                             "B06002_001E", 
                             "B19013_001E", 
                             "B19301_001E"), 
                    region = "county",
                    key = cs_key)
head(acs_2021)

acs_2021 <-
  acs_2021 %>%
  rename(pop = B01001_001E, 
         age = B06002_001E, 
         hh_income = B19013_001E, 
         income = B19301_001E)

acs_2021 <- acs_2021 |> 
        mutate(
        location = paste0(state, county),
         .after = county)
```

-   First, check how many counties aren't matched. Then, create a new data set by joining the two datasets. Keep only counties that appear in both data sets.

**Answer**

Unmatched counties:

647 counties in the covid data set that aren't in acs_2021.

26 counties in acs_2021 dataset that aren't in covid data.

Matched:

6,486 matched pairs at total. There are 807 counties in the combined data set.

```{r}
#| label: acs-match-merge
#| message: false
#| results: hide
covid_both |> 
  distinct(location = geo_value) |> 
  anti_join(acs_2021) 

acs_2021 |> 
  distinct(geo_value = location) |> 
  anti_join(covid_both) 

combined2_full <- inner_join(acs_2021, covid_both, join_by(location == geo_value)) 

combined2_full |> 
  count(location) |> 
  filter(n > 1)
```

-   Compute the mean of the proportion of people with covid-like illness symptoms on April 6, 2020 for counties that have an above average median household income and for those that have an below average median household income. When building your pipe, start with creating the grouping variable and then proceed with the remaining tasks. What conclusions might you draw from this?

**Answer**

Then mean of the proportion of people with covid-like illness symptoms on April 6, 2020 of the above average income group is $0.9624972$, and the mean of the below average income group is $0.9893342$, which is higher than the above one.

This difference indicates that the above average income group is less likely to report illness symptoms than the below average income group.

```{r}
#| label: acs2021-mean-hh

combined_2021 <- combined2_full |> 
  select(location, hh_income, time_value, value.x)

combined_2021 |>
  group_by(above_avg_hh = hh_income > mean(combined_2021$hh_income)) |>
  filter(time_value == as.Date("2020-04-06")) |>
  summarize(mean_prop = mean(value.x))
```

-   Is there a relationship between the median household income and the proportion of people reporting Covid-like illness symptoms? Describe the relationship and use a scatterplot.

**Answer**

The Correlation coefficient is $-0.08024641$, indicating a weak, negative correlation. However, the p-value is 9.678e-11, which is lower than $0.05$, so the correlation is statistically significant. The 95% CI is $[-0.10437997, -0.05601841]$ which does not include zero, further indicating the null hypothesis of no correlation can be rejected.

The scatterplot suggests a weak negative relationship between median household income and the proportion of residents reporting illness symptoms. While there may be a slight trend, the wide scatter and presence of outliers indicate the relationship is inconsistent across counties.

In sum, there is evidence of a significant but weak negative correlation between median household income and COVID-like symptom reporting by county.

```{r}
#| label: cor-acs2021
#| 
cor.test(combined_2021$hh_income, combined_2021$value.x)

ggplot(combined_2021, aes(x = hh_income, y = value.x)) +
  geom_point() +
  geom_smooth(method = "lm",
              se = TRUE,
              color = "blue") +
  labs(x = "Median Household Income", 
       y = "Proportion reporting COVID-like symptoms", 
       title = "COVID-like Illness vs. Median Household Income")
```
